### 2019-12-06 defesa/tex/chapter/definicoes.tex
wip

@@

A function $p_n: \mathbb{C} \mapsto \mathbb{C}$ is a $n$-degree polynomial if it can be written as:
\begin{equation}\label{eq:poly}
p_n(x) = \sum_{k=0}^{n} a_k x^k.
\end{equation}
The famous Abel-Ruffini Theorem (a proof can be seen in \citeonline{skopenkov2015}) states that for polynomials of degree higher than four, there is no closed formula to determine their roots. However, a numerical approach exists for higher-degree polynomials.

 In \citeonline[p.~195]{horn} a theorem is presented which says that for every univariate polynomial, there exists a companion matrix, such that its eigenvalues are the zeros of that polynomial. At first it seems that we are just trading for the same problem as, in general, to find the eigenvalues of a matrix, one needs to determine the roots of that matrix's characteristic polynomial. Nevertheless, it turns out that companion matrices are very convenient when it comes to finding eigenvalues. They are what is called Hessenberg matrices, and QR algorithms can be used to find their eigenvalues (a $\bigO(n^2)$ algorithm that does that can be found in \citeonline{bini:2007}).
 For example, given a polynomial of degree $4$ written as \autoref{eq:poly} its companion matrix is given by:

 \begin{equation}
 \left[\begin{array}{ccccc}
 0 & 1 & 0 & 0 & 0\\
 0 & 0 & 1 & 0 & 0\\
 0 & 0 & 0 & 1 & 0\\
 0 & 0 & 0 & 0 & 1\\
 -a_0 & -a_1 & -a_2 & -a_3 & -a_4
 \end{array}\right].
 \end{equation}

>>>>

\section{Complex numbers}

We introduce here some things about complex numbers that are used later on \autoref{chapter:e3p}.

Any complex number $z = a + bi \in \mathbb{C}$ can be written as follows:
\begin{equation}
z = re^{i\theta},
\end{equation}
where $\theta$ is the polar angle of the vector $(a, b)$, we refer to it as $angle(z)$, and $r$ is the length of this vector, we refer to it as $|z| = \sqrt{a^2 + b^2}$. This is known as the Euler's Formula. If $r=1$, then we say that $z$ is in the unit circle. The complex conjugate is another important operator that we need. Let $z = a + bi \in \mathbb{C}$, then we refer to $\bar{z}$ as the complex conjugate of $z$ and it is defined as:
\begin{equation}
\bar{z} = a - bi.
\end{equation}
Also, it is useful to see that, then $1/z=\bar{z}/|z|$. Two observations about the angles of complex numbers are worth mentioning:
\begin{align}
angle(\bar{z}) = -angle(z)\\
angle(-z) = \pi - angle(z).
\end{align}

@@

A function $p_n: \mathbb{C} \mapsto \mathbb{C}$ is a $n$-degree polynomial if it can be written as:
\begin{equation}\label{eq:poly}
p_n(z) = \sum_{k=0}^{n} a_k z^k.
\end{equation}
The famous Abel-Ruffini Theorem (a proof can be seen in \citeonline{skopenkov2015}) states that for polynomials of degree higher than four, there is no closed formula to determine their roots. However, a numerical approach exists for higher-degree polynomials.

 In \citeonline[p.~195]{horn} a theorem is presented which says that for every univariate polynomial, there exists a companion matrix, such that its eigenvalues are the zeros of that polynomial.  For example, the companion matrix of a degree-$5$ polynomial written as \autoref{eq:poly} is given by:
 \begin{equation}
 \left[\begin{array}{ccccc}
 0 & 1 & 0 & 0 & 0\\
 0 & 0 & 1 & 0 & 0\\
 0 & 0 & 0 & 1 & 0\\
 0 & 0 & 0 & 0 & 1\\
 -\dfrac{a_0}{a_5} & -\dfrac{a_1}{a_5} & -\dfrac{a_2}{a_5} & -\dfrac{a_3}{a_5} & -\dfrac{a_4}{a_5}
 \end{array}\right].
 \end{equation}

 Finding every eigenvalue of a matrix can be done using the QR algorithm, which runs in $\bigO(n^3)$ and uses $\bigO(n^2)$ memory. In \cite[p.~341]{golub} a practical version of it is developed. The idea is to convert the input matrix to the Hessenberg form, after that, the matrix preserves its form through the iterations and under some assumptions, convergence to the eigenvalues is achieved. It is worth mentioning that this approach is actually used in MATLAB and LAPACK, which is the external library used in NumPy for Python.
 Also, in \citeonline{barel} it is pointed that companion matrices are already in the Hessenberg form, making the Hessenberg conversion step unnecessary. This and other properties are used in \citeonline{barel} on the development of a $\bigO(n^2)$ algorithm for the specific case of companion matrices.

### 2019-12-06 defesa/tex/chapter/e3p.tex
wip

As it was pointed out on \autoref{chapter:definitions}, it is possible to find the roots of a univariate polynomial through an algorithm. For that reason, converting $\xi$ somehow into a polynomial can provide a good way to get the solutions of E3PNT.

The first attempt was using the identity $x = \tan{\frac{\theta}{2}}$. This transforms $\xi$ into a polynomial of degree $12$. At first this method seemed to work fine, however, for some instances it did not find roots which were priorly known. This happened probably because of the transformation that was used to convert $\xi$ into a polynomial. It can be very numerically unstable as the tangent function grows to fast for angles that are greater than $\frac{\pi}{4}$. Because of that, depending on the instance, there could be roots very close to $0$ and others that are too far from it.

The second approach is based on a idea published on \citeonline{boyd:2006} which converts a real trigonometric polynomial into a polynomial with a variable $z = e^{i\theta}$. In our case, we replace $\cos{\theta}$ and $\sin{\theta}$ with the following identities:
\begin{align}
\cos{\theta} = \dfrac{e^{i\theta} + e^{-i\theta}}{2}\\
\sin{\theta} = \dfrac{e^{i\theta} - e^{-i\theta}}{2i}.
\end{align}

It is possible to show that with that substitution and changing the variable to $z=e^{it}$, we obtain a function on the form:
\begin{equation}
g(z)=\sum_{k=-6}^6 c_k z^k.
\end{equation}
Then, multiplying $g$ by $z^6 = e^{6i\theta}$, the following polynomial is obtained:

>>>>

On \autoref{chapter:definitions} a brief introduction is given on how to get the roots of a polynomial. For that reason, we discuss two ways of converting $\xi$ into a polynomial. Symbolic computing was used to compute the polynomials, in practice an external Python library called SymPy was utilized (see \citeonline{sympy} for more details).

The first attempt was using the identity $x = \tan{\frac{\theta}{2}}$ from which it is possible to construct a $12$-degree polynomial. At first,  the root-finding algorithm described on \autoref{chapter:definitions} seemed to work fine and return every solution of E3PNT, however, we later found out that for some instances, priorly known roots were not being found. The cause was not for sure identified, but a good guess would be that for angles which are greater than $\frac{\pi}{4}$, $x$ starts growing too rapidly which could lead to numerical instability.

The second approach is based on a idea published on \citeonline{boyd:2006} which uses the identities on \autoref{eq:complex_trig} to convert real trigonometric polynomials into univariate complex polynomials in order to obtain its roots using the companion matrix scheme.
Also, in \citeonline{weidner} it is said that computing the roots of a real trigonometric polynomial through this transformation does not yield loss of accuracy.

\begin{align}\label{eq:complex_trig}
\cos{\theta} = \dfrac{e^{i\theta} + e^{-i\theta}}{2}\\
\sin{\theta} = \dfrac{e^{i\theta} - e^{-i\theta}}{2i}.
\end{align}

It is possible to show that with that substitution and changing the variable to $z=e^{it}$, we obtain the following function $g : \mathbb{S} \mapsto \mathbb{C}$:

\begin{equation}
g(z)=\sum_{k=0}^{12} c_k z^{k-6},
\end{equation}
for some $c_0, \dots, c_{12} \in \mathbb{C}$. It is important to notice that this definition of $g$ does not impose that $z$ stays in the unit circle as any complex number, for that the additional condition: $|z|=1$ would need to be added.

Then, multiplying $g$ by $z^6 = e^{6i\theta}$, the following polynomial is obtained:
